Abstractions and exemplars: the measure noun phrase alternation in German

### 2018-04-09 Latex/sections/4experiments.tex
Basically done and below 13,000 words.

In the two experiments reported in this section, I use probabilities for the alternating constructions calculated for attested material, and I correlate these probabilities with the participants' reactions.
Thus, a direct link can be established between output material found in corpora and the behaviour of linguistic agents.
Both experiments use sentences containing attested MNPs from the corpus sample (embedded into simplified sentences) as stimuli.

Readers might wonder why the stimuli were chosen in such a way.
Alternatively, one could have created stimuli where the features favouring the alternatives were permuted appropriately and thus tested directly.
The answer is related to what was said in Section~\ref{sec:cogocl} about the rich multi-factorial data sets used in corpus studies and the comparatively restricted ones in experiments.
Looking at Table~\ref{tab:variables}, we see one binary factor (\textit{Cardinal}), two three-level factors (\textit{Measurecase} and \textit{Kindgender}) and one five-level factor (\textit{Measureclass}).
Since the dependent variable is binary (\textit{Construction}), permuting these factors alone leads to 180 different possible permutations.
On top of this, the numeric variables \textit{Kindattraction} and \textit{Measureattraction} would also have to be tested at at least two or three different values each.
It is obvious that this is impossible in a controlled experiment, given that participants can usually be exposed to something in the region of one hundred sentences (including many more fillers than target stimuli) within thirty minutes to an hour.
Thus, while the present approach only allows for a global test of the corpus-derived model, this appears to me to be the only feasible way.%
\footnote{Obviously, the experimental data are much too sparse to perform even post-hoc analyses with respect to the single regressors in a statistically and scientifically sound way.}
It is, in the sense of Section~\ref{sec:cogocl}, an optimal synthesis of data-rich multifactorial corpus studies and experimental validation, and it is one that has been used (although rarely; see \citealp[3--4]{DivjakEa2016a}) at least since the seminal \citet{BresnanEa2007} paper.

@@

The first experiment tests preferences for constructions explicitly in a forced choice task.
Participants had to choose between two sentences that differed only in that one contained the \NACa\, and the other one contained the \PGCa.
The analysis compares the probabilities assigned to the stimuli by the corpus-derived model with the frequency with which participants chose the alternants.
There were 24 participants (native speakers of German without reading or writing disabilities) aged 19 to 30 living permanently in Berlin, who were recruited from introductory linguistics courses at Freie Universität Berlin.
Although the experiment was conducted in the last four weeks of their first semester, participants had no deeper explicit knowledge of linguistics, grammar, or experimental methods.
None of them had ever participated in a forced-choice experiment before.
Participation was voluntary but participants received credit in partial fulfillment of course requirements.

As stimuli, attested MNPs from the corpus study were used, but the sentences were simplified to avoid influences from contextual nuisance factors as much as possible.
The approach is also justified because according to the theoretical assessment in Section~\ref{sec:analyses}, the choice of alternants depends mostly on a very local constructional context.
I sampled 16 MNPs from the concordance and made sure that the simplifications and normalisations did not affect any of the regressors used in the corpus study.
In the simplified sentences, the case, number, etc. of the MNP remained the same as in the attested sentence, as did the choice of lexical material within the MNP.
Eight sentences contained masculine or neuter kind nouns, and the other eight contained feminine kind nouns.
Furthermore, in each of the masculine\slash neuter and feminine groups, four sentences originally containing the \NACa\ and four sentences originally containing the \PGCa\ were chosen.
More precisely, the sentences were sampled as \textit{highly typical examples} of \PGCa\ (high probability assigned by GLMM) and \NACa\ (low probability assigned by GLMM), respectively.%
\footnote{Remember from Section~\ref{sec:corpusstudies} that the model predicts the probability that the \PGCa\ is chosen over the \NACa.}
High and low probabilities were defined as the top and bottom 20\% of all probabilities assigned by the GLMM.
Lemmas and feature combinations were made unique within each group whenever possible.
The design is summarised in Table \ref{tab:experiment1:design}.

@@

The final pairs of stimuli were the sentence containing the attested and preferred alternant (according to the corpus GLMM) on the one hand and a modified version containing the dispreferred alternant on the other hand.
They were presented next to each other, and a 20 second time limit for each choice was set.%
\footnote{No participant ever exceeded the time limit.}
The position on the screen (left\slash right) and the order of sentences were randomised for each participant.
As fillers, 23 pairs of sentences exemplifying similar but unrelated alternation phenomena from German morpho-syntax were used.
Thus, participants saw 39 pairs of sentences and 78 sentences in total.
They were instructed to select from each pair of sentences the one that seemed more natural to them in the sense that they would use it rather than the other one.
The experiment was conducted using \textit{PsychoPy} \citep{Peirce2007}.

@@

A multilevel logistic regression was specified with the probability of the \PGCa\ predicted for each sentence by the corpus-based GLMM as the only fixed effect \textit{Modelprediction}.%
\footnote{The document-level variables \textit{Badness} and \textit{Genitives} were set to 0, which is the mean for z-transformed variables.}
A random intercept and slope were added for the individual sentence pair (item) in order to catch idiosyncrasies of single sentences.
Coefficients were estimated with Maximum Likelihood Estimation (\textit{lmer} function from \textit{lme4}).
The number of observations was \textit{n=}384.

@@

In principle, random slope for \textit{Modelprediction} varying by item could also remedy problems with individual stimuli at least partially.
Therefore, a model with random slopes for \textit{Modelprediction} varying by both random effects (\textit{Participant} and \textit{Item}) was specified and estimated.
The random slope for participants was added to comply with \citet[257]{BarrEa2013} who predict ``catastrophically high Type I error rates'' for experimental designs with within-subject manipulations if random effects structures are not kept maximal.
The coefficient of the fixed effect changed noticeably but not enough to change the interpretation (5.408 relative to an intercept of -1.304), and the marginal $R^2_m$ rises to 0.213 ($R^2_c=0.488$).
In line with expectations, the standard deviation in the random slopes for \textit{Item} is high at 5.996.
However, the covariance parameters were estimated at -1.0, which is a clear sign that the variance-covariance matrix could not be estimated successfully.
The same was true for models with only an \textit{Item} and a \textit{Participant} random slope.
This is exactly the kind of model overparametrisation criticised in \citet{BatesEa2015a} and \citet{MatuschekEa2017}.
The available data are insufficient to estimate the parameters of the more complex model with varying slopes.%
\footnote{\citet[1]{BatesEa2015a} state: ``We show that failure to converge typically is not due to a suboptimal estimation algorithm, but is a consequence of attempting to fit a model that is too complex to be properly supported by the data, irrespective of whether estimation is based on maximum likelihood or on Bayesian hierarchical modeling with uninformative or weakly informative priors.
Importantly, even under convergence, overparameterization may lead to uninterpretable models.''
}

@@

The second experiment tests preferences more implicitly.
It is expected that reading less typical alternants in a given context and with given lexical material incurs a processing overhead for the reader (\citealp{Kaiser2013}).
In this section, a self-paced reading experiment is therefore presented.
In a very similar fashion, \cite{DivjakEa2016} apply the self-paced reading paradigm in the validation of corpus-based models.
The analysis compares the corpus-derived probabilities with potential lags in reading time for sentences with the preferred and the non-preferred constructions.

Concretely, the exact same stimuli as in the forced-choice experiments were used.
Each participant read both the 16 sentences with the alternant predicted by the corpus model and the 16 modified sentences with the alternant that the corpus model did not predict.%
\footnote{Notice that lemmas and their frequencies as well as lemma classes are included as regressors in the corpus-based GLMM, and there was consequently no additional controlling of lemma frequencies, etc.}
To minimise repetition effects, the stimuli for each participant were separated into two blocks of 16 targets and 33 fillers per block.
In the experiment, participants first read all sentences from the first block, then all sentences from the second block.
From each target sentence pair, one sentence was assigned to the first block and the other sentence to the other block.
The assignment of members of the individual sentence pairs to the blocks was randomised for each participant individually, as was the order within each block.
The fillers also came in pairs such that the second block exclusively contained sentences to which participants had been exposed in the first block in slightly modified form.
In total, each participant read 98 sentences.
After each sentence, participants had to answer simple (non-metalinguistic) yes-no questions about the previous sentence as distractors.
The distractor questions were different between the first and the second blocks.
There were 38 participants recruited in exactly the same manner as for the experiment reported in Section~\ref{sec:exp:fc}.
None of them had ever participated in any kind of reading experiment, and none of them took part in the first experiment.
The experiment was conducted using \textit{PsychoPy}.

@@

An LMM was specified with the logarithmised residual reading times as the response variable.
The probabilities derived from the corpus GLMM (\textit{Modelprediction}) were added as the main regressor of interest.
It should be remembered that the corpus GLMM predicts the probability of the \PGCa.
As a consequence, the higher the GLMM prediction is, the more typical the sentence is for containing the \PGCa.
It is therefore expected that reading times are higher when the value of \textit{Modelprediction} is higher but the sentence contains the \NACa.
However, when the sentence contains the \PGCa, reading times should be lower when \textit{Modelprediction} is higher.
To account for this, an interaction between \textit{Modelprediction} and \textit{Construction} (levels \textit{PGCadj} and \textit{NACadj}) was added to the model.

@@

The coefficients of determination indicate that there is a noteworthy correlation between the reactions of the subjects and the corpus-derived probabilities (marginal $R^2$) and that there is some between-subject variation (conditional minus marginal $R^2$).

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.5\textwidth]{../R/output/spr_effects}
\caption{Effect plot for the LMM in the self-paced reading experiment: modeling participants' residualised log reading times on the probabilities given by the corpus-based GLMM}
\label{fig:spr:effects}
\end{figure}

>>>>

In this section, the probabilities of the alternants as calculated for attested material are correlated with the reactions of participants in controlled experiments.
Thus, a direct link is established between output material found in corpora and the behaviour of linguistic agents.

The decision to use attested material and perform a global validation of the corpus-based model was not arbitrary.
Alternatively, one could have created stimuli where the features favouring the alternatives were permuted appropriately and thus tested directly.
The answer is related to what was said in Section~\ref{sec:cogocl} about the rich multi-factorial data sets used in corpus studies and the comparatively restricted ones in experiments.
Looking at Table~\ref{tab:variables}, we see one binary factor (\textit{Cardinal}), two three-level factors (\textit{Measurecase} and \textit{Kindgender}) and one five-level factor (\textit{Measureclass}).
Since the dependent variable is binary (\textit{Construction}), permuting these factors alone leads to 180 different possible permutations.
On top of this, the numeric variables \textit{Kindattraction} and \textit{Measureattraction} would also have to be tested at several values.
It is obvious that this is impossible in a controlled experiment, given that a participant can be exposed to no more than roughly one hundred sentences, most of which would be fillers rather than target sentences.
Thus, while the present approach only allows for a global test of the corpus-derived model, this appears to me to be the only feasible way.%
\footnote{For the same reasons, the experimental data are also much too sparse to perform even post-hoc analyses with respect to the single regressors in a sound way.}
It is, in the sense of Section~\ref{sec:cogocl}, an optimal synthesis of data-rich multifactorial corpus studies and experimental validation, and it is one that has been used (although rarely; see \citealp[3--4]{DivjakEa2016a}) at least since the seminal \citet{BresnanEa2007} paper.

@@

The first experiment tests preferences for constructions explicitly in a forced-choice task.
Participants had to choose between two sentences that differed only in that one contained the \NACa\, and the other one contained the \PGCa.
In the analysis, the probabilities assigned to the stimuli by the corpus-derived model were compared to the frequency with which participants chose the alternants.
There participants were 24 native speakers of German without reading or writing disabilities, aged 19 to 30, and living permanently in Berlin.
They were recruited from introductory linguistics courses at Freie Universität Berlin.
Although the experiment was conducted in the last four weeks of their first semester, participants had no deeper explicit knowledge of linguistics, grammar, or experimental methods.
None of them had ever participated in a forced-choice experiment before.
Participation was voluntary, but participants received credit in partial fulfillment of course requirements.

As stimuli, attested MNPs from the corpus study were used, but the sentences were simplified to avoid influences from contextual nuisance factors as much as possible.
16 MNPs were sampled from the concordance, making sure that any simplifications and normalisations did not affect any of the regressors used in the corpus study.
In the simplified sentences, the case, number, etc. of the MNP remained the same as in the attested sentence, as did the choice of lexical material within the MNP.
Eight sentences contained masculine or neuter kind nouns, and the other eight contained feminine kind nouns.
Furthermore, in each of the masculine\slash neuter and feminine groups, four sentences originally containing the \NACa\ and four sentences originally containing the \PGCa\ were chosen.
Moreover, the sentences were sampled as typical examples of \PGCa\ (high probability assigned by GLMM) and \NACa\ (low probability assigned by GLMM), respectively.%
\footnote{Remember from Section~\ref{sec:corpusstudies} that the model predicts the probability that the \PGCa\ is chosen over the \NACa.}
High and low probabilities were defined (roughly) as the top and bottom 20\% of all probabilities assigned by the GLMM.
Lemmas and feature combinations were made unique within each group whenever possible.
For calculating the model predictions, the document-level variables \textit{Badness} and \textit{Genitives} were set to 0, which is the mean for z-transformed variables.
The design is summarised in Table \ref{tab:experiment1:design}.

@@

The final pairs of stimuli were the sentence containing the attested and preferred alternant (according to the corpus GLMM) on the one hand and a modified version containing the dispreferred alternant on the other hand.
They were presented next to each other, and a 20 second time limit for each choice was set.%
\footnote{No participant ever exceeded the time limit.}
The order of sentences was randomised for each participant, and the position of the alternants on the screen (left\slash right) was randomised per participant and sentence.
As fillers, 23 pairs of sentences exemplifying similar but unrelated alternation phenomena from German morpho-syntax were used.
Thus, participants saw 39 pairs of sentences and 78 sentences in total.
They were instructed to select from each pair of sentences the one that seemed more natural to them in the sense that they would use it rather than the other one.
The experiment was conducted using \textit{PsychoPy} \citep{Peirce2007}.

@@

A multilevel logistic regression was specified with the probability of the \PGCa\ predicted for each sentence by the corpus-based GLMM as the only fixed effect \textit{Modelprediction}.
A random intercept was added for the individual sentence pair (\textit{Item}) in order to catch idiosyncrasies of single sentences.
Coefficients were estimated with Maximum Likelihood Estimation (\textit{lmer} function from \textit{lme4}).
The number of observations was \textit{n=}384.

@@

In principle, random slope for \textit{Modelprediction} varying by item could also remedy problems with individual stimuli at least partially.
Therefore, a model with random slopes for \textit{Modelprediction} varying by both random effects (\textit{Participant} and \textit{Item}) was specified and estimated.
The random slope for participants was added to comply with \citet[257]{BarrEa2013} who predict ``catastrophically high Type I error rates'' for experimental designs with within-subject manipulations if random effects structures are not kept maximal.
The coefficient of the fixed effect changed noticeably but not enough to change the interpretation (5.408 relative to an intercept of -1.304), and the marginal $R^2_m$ rises to 0.213 ($R^2_c=0.488$).
In line with expectations, the standard deviation in the random slopes for \textit{Item} is high at 5.996.
However, the covariance parameters were estimated at exactly -1, which is a clear sign that the variance-covariance matrix could not be estimated successfully.
The same was true for models with only an \textit{Item} and a \textit{Participant} random slope.
This is exactly the kind of model overparametrisation criticised in \citet{BatesEa2015a} and \citet{MatuschekEa2017}.
The available data are insufficient to estimate the parameters of the more complex model with varying slopes.%
\footnote{\citet[1]{BatesEa2015a} state: ``We show that failure to converge typically is not due to a suboptimal estimation algorithm, but is a consequence of attempting to fit a model that is too complex to be properly supported by the data, irrespective of whether estimation is based on maximum likelihood or on Bayesian hierarchical modelling with uninformative or weakly informative priors.
Importantly, even under convergence, overparameterization may lead to uninterpretable models.''
}

@@

The second experiment was conducted in a more implicit paradigm.
It is expected that reading less typical alternants in a given context and with given lexical material incurs a processing overhead for the reader (\citealp{Kaiser2013}).
In this section, a self-paced reading experiment is therefore presented.
In a very similar fashion, \cite{DivjakEa2016} apply the self-paced reading paradigm in the validation of corpus-based models.
The analysis compares the corpus-derived probabilities with potential lags in reading time for sentences with the preferred and the non-preferred constructions.

The exact same stimuli as in the forced-choice experiments were used.
Each participant read both the 16 sentences with the alternant predicted by the corpus model and the 16 modified sentences with the alternant that the corpus model did not predict.%
\footnote{Notice that lemmas and their frequencies as well as lemma classes are included as regressors in the corpus-based GLMM, and there was consequently no additional controlling of lemma frequencies, etc.}
To minimise repetition effects, the stimuli for each participant were separated into two blocks of 16 targets and 33 fillers per block.
In the experiment, participants first read all sentences from the first block, then all sentences from the second block.
From each target sentence pair, one sentence was assigned to the first block and the other sentence to the other block.
The assignment of members of the individual sentence pairs to the blocks was randomised for each participant individually, as was the order within each block.
The fillers also came in pairs such that the second block exclusively contained sentences to which participants had been exposed in the first block in slightly modified form.
In total, each participant read 98 sentences.
After each sentence, participants had to answer simple (non-metalinguistic) yes-no questions about the previous sentence as distractors.
The distractor questions were different between the first and the second blocks.
There were 38 participants recruited in exactly the same manner as for the experiment reported in Section~\ref{sec:exp:fc}.
None of them had ever participated in any kind of reading experiment, and none of them took part in the first experiment.
The experiment was conducted using \textit{PsychoPy}.

@@

An LMM was specified with the logarithmised residual reading times as the response variable.
The probabilities derived from the corpus GLMM (\textit{Modelprediction}) were added as the main regressor of interest.
It should be remembered that the higher the GLMM prediction is, the more typical the sentence is for containing the \PGCa.
It is therefore expected that reading times are higher when the value of \textit{Modelprediction} is higher but the sentence contains the \NACa.
However, when the sentence contains the \PGCa, reading times should be lower when \textit{Modelprediction} is higher.
To account for this, an interaction between \textit{Modelprediction} and \textit{Construction} (levels \textit{PGCadj} and \textit{NACadj}) was added to the model.

@@

The coefficients of determination indicate that there is a noteworthy correlation between the reactions of the subjects and the corpus-derived probabilities (marginal $R^2$) and that there is some between-subject variation (conditional minus marginal $R^2$, which is 0.107).

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.5\textwidth]{../R/output/spr_effects}
\caption{Effect plot for the LMM in the self-paced reading experiment: modelling participants' residualised log reading times on the probabilities given by the corpus-based GLMM}
\label{fig:spr:effects}
\end{figure}

### 2018-04-09 Latex/sections/5conclusions.tex
Basically done and below 13,000 words.

The empirical studies in Sections~\ref{sec:corpusstudies} and~\ref{sec:experimental} confirm the theoretical model proposed in Section~\ref{sec:germanmeasurenps}, which predicted prototype effects and exemplar effects to jointly determine which of the two alternating constructions is chosen.
More specifically, the pre-study and the main study provided evidence that the NAC constructions favour more strongly grammaticalised measure nouns, modification by cardinals, and styles where the genitive is underrepresented and the text is less coherently and elaborately written, as captured in the \textit{Badness} variable.%
\footnote{Although, as pointed out in Section~\ref{sec:mainstudyinterpretation}, the \textit{Badness} effect is weak and not detected in the pre-study, probably as a consequence of its weakness.}
The overall convergence between the pre-study and the main study and also the expected result for \textit{Measurecase} in line with previous research \citep{Zimmer2015} -- although it was only a control variable here -- strengthen the validity of this study even without the experimental validation.

While the results do not \textit{prove} -- due to the intrinsic equivalence of prototype and exemplar models with respect to the output they produce -- that the model proposed in Section~\ref{sec:germanmeasurenps} is congruent with any speaker's mental representation, the two attraction features (\textit{Kindattraction} and \textit{Measureattraction}) are virtually impossible to conceive of as prototype effects.
As discussed in Section~\ref{sec:cogocl}, it is unnecessary to follow an extreme route, be it radical prototype theory or radical exemplar theory given recent developments in cognitive science and (less prominently) cognitive linguistics.
While it is surely an intentionally overstated comment, \citet[15]{Kapatsinski2014} even suggests that ``[i]n the extreme, some speakers’ heads could host exemplar models, and some could contain fairly abstract grammars, and the produced output would be essentially identical''.
If this is the case (even to a less extreme degree), corpora only provide (still informative) pooled data averaged over speaker grammars with varying levels of abstraction.
Clearly, more research is required in this direction.

@@

\begin{quote}
  There are now a number of published multivariate models that use data[,] extracted from corpora [\ldots] to predict the choice for one morpheme, lexeme or construction over another.
  However, [\ldots] only a small number of these corpus-based studies have been cross-validated [\ldots].
  Of these cross- validated studies, few have directly evaluated the prediction accuracy of a complex, multivariate corpus-based model on humans using authentic corpus sentences [\ldots].
\end{quote}

The question now arises whether convergence was reached in the present case or not.
The answer is a clear yes.
The predictions made by the corpus-based model were a significant factor, both in the more explicit paradigm (forced choice) and the implicit paradigm (self-paced reading).
This shows that the model indeed predicts the variant which language users expect.
The overall effect as measured in the $R^2$ was weak (roughly 0.2) in both cases, however.
While part of this could be traced back to one suboptimally chosen stimulus, we really need to consider what kind of convergence we expect to see between corpus data and experiments.
The main corpus study had $R^2_m=0.409$ and $R^2_c=0.495$, which is good but not anywhere near a perfect fit.
With the added inter-speaker variability brought into play by the experimental setup (compared to the averaging across thousands of speakers in the corpus study), a perfect fit cannot be expected.
Like many phenomena in German which are often called \textit{Zweifelsfälle} (`cases of doubt') in the traditional literature \citep{Duden09,Klein2009} the MNP alternation is one of the cases where speakers very often have no clear intuition and a lot of free variation seems to be involved.
Rarely do speakers feel that one alternative is clearly odd or bad.
Additionally, cases of doubt involving the genitive are often a matter of fierce normative public debate, and especially the forced choice paradigm does not effectively prevent participants from making normative judgements.
This might even account for the slightly better fit in the self-paced reading experiment, where normative considerations are suppressed.
Thus, the present study shows that what counts as convergence between corpus and experimental results should be gauged considering the nature of the phenomenon at hand, the source of the corpus data, and the experimental paradigm.
Clearly, more case studies using diverse and different corpora are needed, and it should become standard practice to cross-validate them using experimental methods.
Given that a single failure to achieve convergence does not provide conclusive evidence for divergence, many more studies need to be published to the point that meta-analysis becomes possible.%
\footnote{This is even more vital considering the variation in results from experimental work.
See, for example, the impressive list of different reading time results from ten papers on Chinese relative clause processing in \citet[8]{Vasishth2015}.}

On a larger methodological scale, this paper also makes a number of contributions.
The statistical model was a true multilevel model (see Sections~\ref{sec:exemplareffects} and~\ref{sec:mainstudystatisticalmodel}), demonstrating how multilevel modelling helps to specify complex hierarchies of abstract features, item-specific tendencies, and exemplar effects at the level of observations (sentences; first level) and lemmas (second level).
While \cite{Gries2015} still calls mixed models ``underused'' in corpus linguistics, multilevel models are consequently at least equally underused tools.
At the same time, the effects of overparametrisation of mixed models with varying slopes as criticised by \cite{BatesEa2015a} were demonstrated.
Furthermore, fitting an additive model to the reading time data did not improve the fit as there were no non-linearities in the data.
While \cite{DivjakEa2016} also use attested sentences, they find that an additive model helped to deal with non-linearities.
This is clearly another area where only more studies can lead to clarification.
Finally, much like \cite{Dabrowska2014} found that speakers' knowledge of collocations was not matched by a set of standard measures of collocation strength extracted from corpora, a simple quotient based on raw frequencies for the attraction strength performed much better in the present study compared to collexeme strength (see Section~\ref{sec:corpusstudies}).
While this does not allow the conclusion that collexeme strength has no cognitive reality, it might indicate that for measuring attraction effects exerted by non-alternating constructions on alternating constructions, other measures might be more appropriate.

Besides the methodological aspects mentioned above, future work could extend the validity of the results presented here through a more in-depth look at stylistic or register effects, for example using the new Biber-style annotations \citep{Biber1988} which will be released with a new version of the DECOW corpus soon according to its creators.
Also, regional variation as a potentially influencing factor was already mentioned in Section~\ref{sec:prototypeeffects}.
As it would be infeasible to examine this with existing corpora, experimental work with participants from various regions might be the ideal approach.
A reviewer also pointed out that strongly lexicalised adjective-noun combinations like \textit{schwarzer Tee} `black tea' might have a tendency to occur in the \NACa\ because they have more compound-like qualities, blocking strong case inflection in between them.
While this potential effect was impossible to incorporate into the present study, it could be integrated into future research on the phenomenon.

>>>>

The studies reported in Sections~\ref{sec:corpusstudies} and~\ref{sec:experimental} have confirmed the theoretical model proposed in Section~\ref{sec:germanmeasurenps}, which predicted prototype effects and exemplar effects to jointly determine which of the two alternating constructions is chosen.
More specifically, the pre-study and the main study provided evidence that the NAC constructions favour more strongly grammaticalised measure nouns, modification by cardinals, and styles where the genitive is underrepresented and the text is less coherently and elaborately written.
The overall convergence between the pre-study and the main study and the convergence between corpus data and the experimental cross-validation strengthen the validity of this study.
Also, the fact that the result for \textit{Measurecase} (although it was only a control variable in the present context) converged with previous research \citep{Zimmer2015} lends credibility to the results.

Many relevant conclusions can be drawn based on the research presented here.
Although direct inferences from corpus data to mental representations are problematic (see Section~\ref{sec:cogocl}), the two attraction features (\textit{Kindattraction} and \textit{Measureattraction}) are virtually impossible to conceive of as prototype effects and thus favour an exemplar view or a mixed exemplar-prototype view.
As discussed in Section~\ref{sec:cogocl}, it is unnecessary to follow an extreme route, be it radical prototype theory or radical exemplar theory, given recent developments in cognitive science and cognitive linguistics.
While it is surely an intentionally overstated comment, \citet[15]{Kapatsinski2014} even suggests that ``[i]n the extreme, some speakers’ heads could host exemplar models, and some could contain fairly abstract grammars, and the produced output would be essentially identical''.
If this is the case (even to a less extreme degree), corpora only provide pooled data averaged over speaker grammars with varying levels of abstraction.
Clearly, more research is required in this direction.

@@

\begin{quote}
  There are now a number of published multivariate models that use data[,] extracted from corpora [\ldots] to predict the choice for one morpheme, lexeme or construction over another.
  However, [\ldots] only a small number of these corpus-based studies have been cross-validated [\ldots].
  Of these cross-validated studies, few have directly evaluated the prediction accuracy of a complex, multivariate corpus-based model on humans using authentic corpus sentences [\ldots].
\end{quote}

The question now arises whether convergence was reached in the present case or not.
The answer is a clear yes.
The predictions made by the corpus-based model were a significant factor, both in the more explicit paradigm (forced choice) and the implicit paradigm (self-paced reading).
This shows that the model indeed partially predicts the variant which language users expect.
The overall effect as measured in the $R^2$ was weak (roughly 0.2) in both cases, however.
While part of this could be traced back to one suboptimally chosen stimulus, we really need to consider what kind of convergence we expect to see between corpus data and experiments.
The main corpus study had $R^2_m=0.409$ and $R^2_c=0.495$, which is good but not anywhere near a perfect fit.
With the added inter-speaker variability brought into play by the experimental setup (compared to the averaging across thousands of speakers in the corpus study), a perfect fit cannot be expected.
Like many phenomena in German which are often called \textit{Zweifelsfälle} (`cases of doubt') in the traditional literature \citep{Duden09,Klein2009} the MNP alternation is a case where speakers often have no clear intuition, and a lot of free variation seems to be involved.
Rarely do speakers feel that one alternative is clearly odd or bad.
Additionally, cases of doubt involving the genitive are often a matter of fierce normative public debate, and especially the forced choice paradigm does not effectively prevent participants from making normative judgements.
This might even account for the slightly better fit in the self-paced reading experiment, where normative considerations are suppressed.
Thus, the present study shows that what counts as convergence between corpus and experimental results should be gauged considering the nature of the phenomenon at hand, the source of the corpus data, and the experimental paradigm.
Clearly, more case studies using diverse and different corpora are needed, and it should become standard practice to cross-validate them using experimental methods.
Given that a single failure to achieve convergence does not provide conclusive evidence for divergence, many more studies need to be published to the point that meta-analysis becomes possible.%
\footnote{This is even more vital considering the variation in results from experimental work.
See, for example, the impressive list of different reading time results from ten papers on Chinese relative clause processing in \citet[8]{Vasishth2015}.}

On a larger methodological scale, this paper also makes a number of contributions.
The statistical model was a true multilevel model (see Sections~\ref{sec:exemplareffects} and~\ref{sec:mainstudystatisticalmodel}), demonstrating how multilevel modelling helps to specify complex hierarchies of abstract features, item-specific tendencies, and exemplar effects at the level of observations (sentences; first level) and lemmas (second level).
\cite{Gries2015} still calls mixed models ``underused'' in corpus linguistics, and multilevel models are consequently also an underused tools.
At the same time, the effects of overparametrisation of mixed models with varying slopes as criticised by \cite{BatesEa2015a} were demonstrated.
Furthermore, fitting an additive model to the reading time data did not improve the fit as there were no non-linearities in the data.
While \cite{DivjakEa2016} also use attested sentences, they find that an additive model helped to deal with non-linearities.
This is clearly another area where only more studies can lead to clarification.
Finally, much like \cite{Dabrowska2014} found that speakers' knowledge of collocations was not matched by a set of standard measures of collocation strength extracted from corpora, a simple quotient based on raw frequencies for the attraction strength performed much better in the present study compared to collexeme strength in the form of logarithmised Fisher p-values (see Section~\ref{sec:corpusstudies}).
While this does not allow the conclusion that collexeme strength has no cognitive reality, it might indicate that for measuring attraction effects exerted by non-alternating constructions on alternating constructions, taking the marginals into account might not be crucial.%
\footnote{However, many different measures have been proposed within the collostructional framework (see \citealp{Gries2015b} for an overview of them and a defense of the framework).
Future research might show significant differences between them for the problem at hand.}

Besides the methodological aspects mentioned above, future work could extend the validity of the results presented here through a more in-depth look at stylistic or register effects, for example using the new Biber-style annotations \citep{Biber1988} which will be released with a new version of the DECOW corpus soon according to its creators.
A reviewer also pointed out that strongly lexicalised adjective-noun combinations like \textit{schwarzer Tee} `black tea' might have a tendency to occur in the \NACa\ because they have more compound-like qualities, blocking strong case inflection in between them.
While this potential effect was impossible to incorporate into the present study, it could be integrated into future research on the phenomenon.
